# 6 - SDD results
## setup
```{r}
library(tidyverse)
library(ggmap)
library(janitor)
```
## Visualizing the joined data
### maps
Let's make a map for the joined data.
```{r}
ggmap(map_area) +
  # geom_point(geno_coords, mapping=aes(X, Y), size = 0.5) +
  geom_segment(structure_pedigree,
               mapping=aes(x=parent_x, y=parent_y, xend=offspring_x, yend=offspring_y, color=off_percent_brev),
               arrow = arrow(
                 length = unit(0.01, "npc"),
                 type = "closed",
                 angle = 20),
               size=0.2) +
  labs(color='Offspring % brevifolia')

?labs
```
The arrows point from the parent to the offspring, with darker colors signifying the offspring genetically is more like *Y. jaegeriana* while light blue signifies *Y. brevifolia*.

With this map, we clearly see an outlier spanning nearly the entire length of our range of sample collection. We will be analyzing our results with and without this outlier.

Looking at this data, it seems more paternal relationships were inferred for *Y. brevifolia* than *Y. jaegeriana*. Let's explore that further.

Lets take a look at the dispersal distance compared to % brev.
```{r}
structure_pedigree %>%
  ggplot(aes(x=off_percent_brev, y=distance)) +
  geom_point() +
  theme_classic()
```
We have a wide range of brevifolia, jaegeriana, and hybrids represented in our data. We will be analyzing this categorically via thresholding our data.

## Thresholding data
For purposes of our thresholds, we will be arbitrarily identifying any offspring with % brev lesser than 0.25 as *y. Jaegeriana*, any as greater than 0.75 as *y. Brevifolia*, and those in between  as hybrids.
```{r}
species_sdd <- structure_pedigree %>% 
  mutate(
    species = as.factor(case_when(
      off_percent_brev < 0.25 ~ 'y. Jaegeriana',
      off_percent_brev > 0.75 ~ 'y. Brevifolia',
      .default = 'hybrid'
    ))
  )

summary(species_sdd$species)

species_sdd %>% 
  group_by(species) %>% 
  summarize(mean=mean(distance), sd=sd(distance))
```


```{r}
aov_species_sdd = aov(distance ~ species, data=species_sdd)

summary(aov_species_sdd)
TukeyHSD(aov_species_sdd)
```
[TODO: update this when I have the rest of the data for the analysis]




and then look at it without the outlier
```{r}
structure_pedigree_no_outlier = structure_pedigree %>% 
  filter(distance < 10000)

structure_pedigree_no_outlier %>%
  ggplot(aes(x=off_percent_brev, y=distance)) +
  geom_point() +
  theme_classic()
```





```{r}
trees_distances %>%
  ggplot(aes(x = distance,y=after_stat(density))) +
  geom_histogram(fill="thistle2", binwidth=200) +
  geom_density(color = "red") +
  theme_classic() +
  labs(x='Distance (m)', y='Density', title='Distribution of estimated dispersal distance')+
  facet_wrap(~)
```
No dice on this type.. will have to keep looking


```{r}
trees_distances %>% filter(distance<9000) %>% summary()
```

