# 1 - initial data reading/setup

## setup
setting up required libraries
```{r}
library(tidyverse)
library(janitor)
library(ggmap)
library(spdep)
# library(spdep) [TODO: isn't installing]
```

## Reading in data
For spatial autocorrelation, we're going to analyze a few things. One of which is morphology data, which we haven't read in yet. This is in 'data-files/Morphology.csv'
[TODO: make sure this is the most complete version of our morphology data]

```{r}
morphology <- read_csv('data-files/Morphology.csv') %>% clean_names() %>% unique()
head(morphology)
summary(morphology)
```
Now let's join this to our coordinate data. There's need to connect this to genotype data, as this is separate analysis.
```{r}
morph_coords = inner_join(morphology, coords, by=c('tree'='Tree')) %>% select(tree, X, Y, height)
summary(morph_coords)
```
We're left with 1096 trees:

```{r}
ggmap(map_area) +
  # geom_point(geno_coords, mapping=aes(X, Y), size = 0.5) +
  geom_point(morph_coords,
               mapping=aes(x=X, y=Y),
               size=.5)
```
need to make gemoetric point coordinates for spdep Moran's I functions
```{r}
sf_morph_coords <- morph_coords %>%
  mutate(long=X, lat=Y) %>% 
  st_as_sf(coords = c("X", "Y"), crs=4326)

# 4326 is the code for gps lat/long

```


Now we have to assign our neighbors -- we are doing this distance based -- all within 10m will qualify
```{r}
neighbors <- dnearneigh(x=sf_morph_coords, d1=0, d2=10)
```



Now add weights:
```{r}
weighted <- nb2listw(neighbors, style="W", zero.policy=F)
```


Now we get global Moran's I:
```{r}
MC  <-  moran.mc(sf_morph_coords$height, weighted, nsim=5000, zero.policy=F)
```

```{r}
plot(MC, main="", las=1)
```

```{r}
MC
```
Now local Moran's I:
```{r}
MCi <- localmoran_perm(sf_morph_coords$height, weighted, nsim = 9999)
MCi.df <- as.data.frame(MCi)
```

Now let's add those p values to our data:
```{r}
sf_morph_coords$p <- MCi.df$`Pr(folded) Sim`
summary(sf_morph_coords$p)
```

[todo: Note that the localmoran_perm function generates two different p-values: MCi.df$`Pr(z != E(Ii)) Sim` and MCi.df$`Pr(folded) Sim`. The former is for a two sided test (alternative = "two.sided") and the latter is a “folded” p-value for a one-sided test.]


Now lets map our p values:
```{r}
ggmap(map_area) +
  # geom_point(geno_coords, mapping=aes(X, Y), size = 0.5) +
  geom_point(sf_morph_coords,
               mapping=aes(x=long, y=lat, color=p),
               size=.5)
```



```{r}
ggmap(map_area) +
  # geom_point(geno_coords, mapping=aes(X, Y), size = 0.5) +
  geom_point(sf_morph_coords,
               mapping=aes(x=long, y=lat, color=height),
               size=.5)
```

```{r}
ggmap(map_area) +
  # geom_point(geno_coords, mapping=aes(X, Y), size = 0.5) +
  stat_summary_2d(sf_morph_coords,
               mapping=aes(x=long, y=lat, z=height))
```
```{r}
sf_morph_coords %>% summary()
```


```{r}
sf_morph_coords$Ii <- hotspot(MCi, Prname="Pr(folded) Sim", cutoff = 0.05, p.adjust = "none")

# Replace NA with ">0.05". This requires that the Ii factor be re-leveled
sf_morph_coords$Ii <- factor(sf_morph_coords$Ii, levels=c("High-High","Low-Low", "Low-High", "High-Low", ">0.05"))
sf_morph_coords$Ii[is.na(sf_morph_coords$Ii)] <- ">0.05"
```

```{r}
spatial_species = inner_join(sf_morph_coords%>%select(-geometry), species_sdd, by=c('tree' ='offspring_id'))
summary(species_sdd)


ggmap(map_area) +
  # geom_point(geno_coords, mapping=aes(X, Y), size = 0.5) +
  geom_point(spatial_species,
               mapping=aes(x=long, y=lat, color=Ii),
               size=.5)+
  facet_grid(~species)
```

```{r}
sf_morph_coords %>% 
  group_by(Ii) %>% 
  summarize(mean(height))
```

