---
title: "paired ttest min-max sire-dam"
output: html_document
date: "2024-10-15"
---

```{r setup, include=FALSE}
library(tidyverse)
```

## filtering pedigree

```{r}
pedigree

two_parents = final %>% 
  filter(!is.na(dam)) %>% 
  gather(key = parent_type, value=parent_id, sire, dam)


two_parents_meters = inner_join(distances_m, two_parents,
                                by = c("offspring_id" = "id", "parent_id" = "parent_id"))
  
  
?inner_join

two_parents_meters
```
## okay now to get min and max
```{r}
max_min_distance = two_parents_meters %>%
  group_by(, offspring_id) %>% 
  summarize(Max = max(distance),
            Min = min(distance))

max_min_distance %>% summary()
max_min_distance %>% arrange(desc(Min))
```
## graphing it
```{r}
max_min_distance %>%
  gather(maxmin, distance, -offspring_id) %>% 
  ggplot(aes(x = distance)) +
  geom_histogram(fill="thistle2") +
  theme_classic() +
  facet_wrap(~maxmin) +
  labs(x='Distance (m)', y='Count')
```


## ttest time
```{r}
t.test(max_min_distance$Max, max_min_distance$Min, paired=T)


max_min_distance %>%
  gather(key=group, value=distance, Max, Min) %>%
  group_by(group) %>% 
  summarize(mean = mean(distance), sd = sd(distance), median = median(distance))

max_min_distance %>%
  gather(key=group, value=distance, Max, Min) %>%
  ggplot(mapping=aes(x=distance, fill=group)) +
  geom_histogram() +
  facet_wrap(~group) +
  theme_classic()
```
The maximum and minimum distances were significantly different, with maximum distances being an average of 1276.8 meters greater (paired t-test, t=2.5385, df=33, p < 0.05).

Both groups indicated dispersal distances greater than 

However, this observed difference could result from false paternal relationships, and shorter relationships may be more likely to be correctly inferred may result in




