# Structure analysis
## setup
```{r}
library(tidyverse)
library(ggmap)
library(janitor)
```

Using data from 'distance_franz_pedigree.R'

Will need to join in trees_distances (output data with all the info from the file)

## read in the data
```{r}
structure_assignments <- read_csv('data-files/USGS structure assignments.csv') %>%
  clean_names()
```
## cleaning, exploring, joining with pedigree + distance/coord data
lets start by seeing what each has, and what we should cut from the structure data
```{r}
summary(trees_distances)

summary(structure_assignments)
```
think "Pop" is categorical to show what sort of brev/jaeg/hybrid/backcross it is.. need to ask for more specifics for what this is/investigate it..
either way, should convert it to factor..

also worth noting that not all of this data is from tikaboo valley, so we should filter those out at some point

not sure what to do with the missing category, so let's leave it in for now in case we need to filter based off that later

for now, let's cut out the run 1 and 2, assuming that ave is what we want to work with pending discussion with Chris
```{r}
structure = structure_assignments %>%
  select(-numb, -percent_brev_run1, -percent_brev_run2) %>%
  mutate(pop = as.factor(pop))

summary(structure)
```
lets take a look into changes in trends based off of pop:
```{r}
structure %>%
  group_by(pop) %>% 
  summarize(
    mean_percent_brev = mean(percent_brev_ave),
    n = n()
  )
```
Looks like 2/3/4 are distinct brev populations, 5 is tikaboo?
Ω
```{r}
structure %>% filter(pop == 5)

# and check by filtering for TV ids to see if they match up..

# getting first two characters
structure %>% filter(substr(tree, 1, 2) == "TV") %>% summary()
```
Looks like some from TV are in population 2.. will have to see if that persists when we filter with pedigree data

My bet is the TV which are in pop 2 are part of the TV groups that have other letters after TV in the id...
```{r}
structure %>% filter(substr(tree, 1, 2) == "TV") %>% filter(pop == 2)
```
yep.. these aren't in the pedigree data but should ask Chris about these out of curiosity.. so we can cut out the population when it's time to join

### Let's graph % brev across groups.. I'm curious
```{r}
structure %>%
  ggplot(aes(x=percent_brev_ave))+
  geom_histogram()+
  facet_wrap(~pop)+
  theme_classic()
  
```
Cool cool.. let's not get carried away with this analysis before we join in the pedigree data..

### joining pedigree data
double checking to see if any of the pedigree data connects to multiple parents since I don't remember...
```{r}
trees_distances %>% 
  mutate(offspring_id = as.factor(offspring_id)) %>%
  group_by(offspring_id) %>% 
  summarize(n = n()) %>%
  summary()
```
none of them are.. but if it did, then there would potentially be more reason to prioritize the inclusion of the parental data..  

for now let's just analyze the offspring data

```{r}
# inner join to cut out the 
structure_pedigree = inner_join(trees_distances, structure, by=join_by('offspring_id'=='tree')) %>%
  select(-pop) %>% 
  mutate(off_percent_brev=percent_brev_ave)

summary(structure_pedigree)
```
TODO: add in parent percentage
```{r}
#todo: add this in properly
#inner_join(structure_pedigree, structure, by=join_by('parent_id'=='tree'))

#structure_pedigree = inner_join(structure_pedigree, structure, by=join_by('parent_id'=='tree')) %>%
#  mutate(par_percent_brev=percent_brev_ave) %>% 
#  select(-pop) # all are in pop 5


#summary(structure_pedigree)
```
## visualizing the joined data
### maps
```{r}
ggmap(franz_map_area) +
  # geom_point(geno_coords, mapping=aes(X, Y), size = 0.5) +
  geom_segment(structure_pedigree,
               mapping=aes(x=parent_x, y=parent_y, xend=offspring_x, yend=offspring_y, color=off_percent_brev),
               arrow = arrow(
                 length = unit(0.01, "npc"),
                 type = "closed",
                 angle = 20),
               size=1)
```
based off of this map.. it looks like we have a lot of brevifolia.. not a lot of jaegeriana..  
let's explore the data more. 


### exploratory graphs
lets take a look at the dispersal distance compared to percent brev
```{r}
structure_pedigree %>%
  ggplot(aes(x=off_percent_brev, y=distance)) +
  geom_point() +
  theme_classic()
  
```
## TODO: THRESHOLD AT .25 and .75 and anova


# additional test -- look at age across spatial distribution: spatial autocorrelation
in order of priority:
*anova comparison between tree species
**paired ttest of max and min distances
***spatial statistics looking at height of the trees:
- to consider down the road: filter this off of STRUCTURE data?


## THRESHOLDING NOW
```{r}
species_sdd <- structure_pedigree %>% 
  mutate(
    species = case_when(
      off_percent_brev < 0.25 ~ 'y. Jaegeriana',
      off_percent_brev >0.75 ~ 'y. Brevifolia',
      .default = 'hybrid'
    )
  )


aov_species_sdd = aov(distance ~ species, data=species_sdd)

summary(aov_species_sdd)
TukeyHSD(aov_species_sdd)
```




and then look at it without the outlier
```{r}
structure_pedigree_no_outlier = structure_pedigree %>% 
  filter(distance < 5000)

structure_pedigree_no_outlier %>%
  ggplot(aes(x=off_percent_brev, y=distance)) +
  geom_point() +
  theme_classic()
```
Moran's I might be good for the stastical analysis??

Need to check if distribution is normal enough for that/no big outliers.. putting the code from the distribution graph here

```{r}
trees_distances %>%
  ggplot(aes(x = distance,y=after_stat(density))) +
  geom_histogram(fill="thistle2", binwidth=200) +
  geom_density(color = "red") +
  theme_classic() +
  labs(x='Distance (m)', y='Density', title='Distribution of estimated dispersal distance')
```
No dice on this type.. will have to keep looking


```{r}
trees_distances %>% filter(distance<9000) %>% summary()
```

